### 2018-03-10 paper/background.tex
Polish S2

Another distinguishing feature of \Excel is \emph{self-tracking}. Most build
systems only track changes of inputs and intermediate results, but \Excel can
also track changes in the tasks themselves: if a formula is modified, \Excel
will recompute it and propagate the changes. Self-tracking is uncommon in
software build systems, where one often needs to manually initiate a full
rebuild even if just a single task has changed. We further discuss
self-tracking in~\S\ref{sec-tracking-aspects}.

\begin{figure}[h]
\vspace{-1mm}
\begin{subfigure}[b]{0.90\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/shake-example.pdf}}
\caption{Dependency graph produced after the previous build.}
\end{subfigure}
\begin{subfigure}[b]{0.90\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/shake-example-rebuild.pdf}}
\caption{Since \cmd{docs.txt} has changed, we rebuild \cmd{release.txt} and
\cmd{release.tar}, discovering a new dependency.}
\end{subfigure}
\vspace{-1mm}
\caption{Dynamic dependencies example: create \cmd{README} and add it to the
list of release documents \cmd{docs.txt}.\label{fig-shake}}
\vspace{-2mm}
\end{figure}

@@

\Shake's implementation is different from both \Make and \Excel in two aspects.
First, it uses the dependency from the previous build to detect which files need
to be rebuilt. This is a powerful idea with a long history of
use in \emph{incremental}~\cite{demers1981incremental},
\emph{adaptive}~\cite{acar2002adaptive}, and
\emph{self-adjusting}~\cite{acar2007selfadjusting} \emph{computations} (see a
discussion in related work~\S\ref{sec-related}). Second, instead of abandoning
and deferring the execution of tasks whose newly discovered dependencies have
not yet been built (as \Excel does), \Shake \emph{pauses} their execution until
the dependencies are brought up-to-date. We refer to this build algorithm as
\emph{recursive}.

\begin{figure}[h]
\centerline{\includegraphics[scale=0.28]{fig/make-example-cutoff.pdf}}
\vspace{-2mm}
\caption{An early cutoff example: modify the source \cmd{main.c} by adding a
comment. The rebuild is stopped after detecting that \cmd{main.o} is unchanged,
which indicates that \cmd{main.exe} does not need to be rebuilt.\label{fig-cutoff}}
\end{figure}

@@

\Bazel is one of the first examples of openly available cloud build systems.
Like \Make, it does not support dynamic dependencies and can therefore benefit
from the simplicity of building tasks in the statically known topological order.
It is minimal and supports the early cutoff optimisation.

To support cloud builds, \Bazel maintains a conventional \emph{content-addressable
cache}, which can be used to fetch a previously built file given the hash of its
content, as well as \emph{dependency graphs from all previous builds}, annotated
with observed file hashes\footnote{Here we ignore the issue of limited cloud
storage resources for the sake of simplicity; in practice, old entries are
regularly evicted from the storage, as further discussed
in~\S\ref{sec-cloud-aspects}.}. The latter allows to bypass the execution of
a task, by predicting the hash of its result from the hashes of the dependencies,
and subsequently fetch the result from the cache. A concrete implementation is
provided in~\S\ref{sec-implementations}.

@@

We summarise differences between four discussed build systems in
Table~\ref{tab-summary}. The `Metadata' column refers to the \emph{auxiliary
build information} persistently stored between builds by all build systems:
\begin{itemize}
    \item \Make stores file modification times, or rather, it relies on the file
    system to do that. In principle a single \emph{dirty bit} per file is
    sufficient, as will be demonstrated in~\S\ref{sec-implementations}.
    \item \Excel stores a dirty bit per cell as well as the calculation chain.
    \item \Shake stores the dependency graph discovered in the previous build,
    annotated with file content hashes for efficient checking of file changes.
    \item \Bazel stores all dependency graphs discovered in previous builds
    annotated with file hashes, and the content-addressable cache.
\end{itemize}

\begin{table}[h]
\smaller
\centering
\begin{tabular}{l||l|l||l|c|c|c}
\hline
$\!$Build system$\!$& Metadata                       & Algorithm    & Dependencies & Minimal & Cutoff & Cloud$\!$\\\hline
$\!$\Make       $\!$& File modification times        & Topological  & Static       & Yes     & No     & No   $\!$\\
$\!$\Excel      $\!$& Dirty cells, calculation chain & Reordering   & Dynamic      & No      & No     & No   $\!$\\
$\!$\Shake      $\!$& Last dependency graph          & Recursive    & Dynamic      & Yes     & Yes    & No   $\!$\\
$\!$\Bazel      $\!$& All dependency graphs, cache   & Topological  & Static       & Yes     & Yes    & Yes  $\!$\\\hline
\hline
\end{tabular}
\vspace{2mm}
\caption{Summary of build system differences.\label{tab-summary}}
\vspace{-7mm}
\end{table}

In this paper we show which build system properties are consequences of specific
implementation choices (metadata and algorithm), and how one can obtain new
build systems with desired properties by recombining parts of existing
implementations. As a compelling example, we demonstrate how to combine
advantages of \Shake and \Bazel in a cloud build system with dynamic dependencies.

>>>>

Another distinguishing feature of \Excel is \emph{self-tracking}. Most build
systems only track changes of inputs and intermediate results, but \Excel can
also track changes in the tasks themselves: if a formula is modified, \Excel
will recompute it and propagate the changes. Self-tracking is uncommon in
software build systems, where one often needs to manually initiate a full
rebuild even if just a single build task has changed. We further discuss
self-tracking in~\S\ref{sec-tracking-aspects}.

\begin{figure}[h]
\vspace{1mm}
\begin{subfigure}[b]{0.90\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/shake-example.pdf}}
\vspace{-1mm}
\caption{Dependency graph produced after the previous build.}
\vspace{1mm}
\end{subfigure}
\vspace{1mm}
\begin{subfigure}[b]{0.90\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/shake-example-rebuild.pdf}}
\vspace{-1mm}
\caption{Since \cmd{docs.txt} has changed, we rebuild \cmd{release.txt} and
\cmd{release.tar}, discovering a new dependency.}
\end{subfigure}
\vspace{-2mm}
\caption{Dynamic dependencies example: create \cmd{README} and add it to the
list of release documents \cmd{docs.txt}.\label{fig-shake}}
\vspace{-2mm}
\end{figure}

@@

\Shake's implementation is different from both \Make and \Excel in two aspects.
First, it uses the dependency graph from the previous build to decide which
files need to be rebuilt. This idea has a long history, going back to
\emph{incremental}~\cite{demers1981incremental},
\emph{adaptive}~\cite{acar2002adaptive}, and
\emph{self-adjusting}~\cite{acar2007selfadjusting} \emph{computations} (see a
discussion in related work~\S\ref{sec-related}). Second, instead of abandoning
and deferring the execution of tasks whose newly discovered dependencies have
not yet been built (as \Excel does), \Shake \emph{pauses} their execution until
the dependencies are brought up-to-date. We refer to this build algorithm as
\emph{recursive}.

\begin{figure}[h]
\centerline{\includegraphics[scale=0.28]{fig/shake-example-cutoff.pdf}}
\vspace{-2mm}
\caption{An early cutoff example: if a comment is added to \cmd{main.c}, the
rebuild is stopped after detecting that \cmd{main.o} is unchanged, since this
indicates that \cmd{main.exe} and its dependents do not need to be
rebuilt.\label{fig-cutoff}}
\end{figure}

@@

\Bazel is one of the first examples of openly available cloud build systems.
Like \Make, it does not support dynamic dependencies and can therefore benefit
from the simplicity of building tasks in a statically known topological order.
It is minimal and supports the early cutoff optimisation.

To support cloud builds, \Bazel maintains a conventional \emph{content-addressable
cache}, which can be used to fetch a previously built file given the hash of its
content, as well as \emph{dependency graphs from all previous builds}, annotated
with observed file hashes\footnote{Here we ignore the issue of limited cloud
storage resources for the sake of simplicity; in practice, old entries are
regularly evicted from the storage, as further discussed
in~\S\ref{sec-cloud-aspects}.}. The latter allows to bypass the execution of
a task, by predicting the hash of the result from the hashes of its dependencies,
and subsequently fetching the result from the cache. A concrete implementation is
provided in~\S\ref{sec-implementations}.

@@

We summarise differences between four discussed build systems in
Table~\ref{tab-summary}. The `Metadata' column refers to the \emph{auxiliary
build information} persistently stored between builds by all build systems:
\begin{itemize}
    \item \Make stores file modification times, or rather, it relies on the file
    system to do that. In principle a single \emph{dirty bit} per file is
    sufficient, as will be demonstrated in~\S\ref{sec-implementations}.
    \item \Excel stores one dirty bit per cell and the calculation chain from
    the previous build.
    \item \Shake stores the dependency graph discovered in the previous build,
    annotated with file content hashes for efficient checking of file changes.
    \item \Bazel stores all dependency graphs discovered in previous builds
    annotated with file hashes, and the content-addressable cache.
\end{itemize}

\begin{table}[h]
\smaller
\centering
\begin{tabular}{l||l|l||l|c|c|c}
\hline
$\!$Build system$\!$& Metadata                       & Algorithm    & Dependencies & Minimal & Cutoff & Cloud$\!$\\\hline
$\!$\Make       $\!$& File modification times        & Topological  & Static       & Yes     & No     & No   $\!$\\
$\!$\Excel      $\!$& Dirty cells, calculation chain & Reordering   & Dynamic      & No      & No     & No   $\!$\\
$\!$\Shake      $\!$& Previous dependency graph      & Recursive    & Dynamic      & Yes     & Yes    & No   $\!$\\
$\!$\Bazel      $\!$& All dependency graphs, cache   & Topological  & Static       & Yes     & Yes    & Yes  $\!$\\\hline
\hline
\end{tabular}
\vspace{2mm}
\caption{Summary of build system differences.\label{tab-summary}}
\vspace{-7mm}
\end{table}

In this paper we elucidate which build system properties are consequences of
specific implementation choices (metadata and algorithm), and how one can obtain
new build systems with desired properties by recombining parts of existing
implementations. As a compelling example, we demonstrate how to combine
advantages of \Shake and \Bazel in a cloud build system with dynamic
dependencies, see~\S\ref{sec-implementation-cloud-shake}.
